---
title: "Chapter5"
author: "SebastiánAyala"
date: "9/12/2021"
output: html_document
---

```{r libraries and python version, echo=FALSE, include=FALSE}
library(igraph)
library(ggplot2)
library(reticulate)
library(readr)
library(dplyr)
```

```{python packages, echo=FALSE}
# Import Python packages 
import networkx as nx
import matplotlib.pyplot as plt
import pandas as pd
import numpy as np 
```
# Exercises chapter 5

## 1. Calculate the adjacency matrix, the stochastic adjacency matrix, and the graph Laplacian for this [network](https://www.networkatlas.eu/exercises/5/1/data.txt).

### Python code from [book solutions](https://www.networkatlas.eu/exercise.htm)

```{python exercise 1, echo=FALSE}
# Load the data
G = nx.read_edgelist("../Data/ch5/edge_list_ch5.txt")

# Adjacency matrix
# If this give memory error, try "to_scipy_sparse_matrix"
adjmat = nx.to_numpy_matrix(G)

# Stochastic adjacency
adjmat_stoc = adjmat / adjmat.sum(axis = 1)

# Graph Laplacian
laplacian = nx.laplacian_matrix(G)

```

### R code

```{r excercise 1}
# Load edge list 
#g_edge_list <- read_graph("../Data/ch5/edge_list_ch5.txt", format ="edgelist")
#g_edge_list <- delete.vertices(g_edge_list, "1")
edge_list <- data.frame(read.table("../Data/ch5/edge_list_ch5.txt", col.names = c("V1", "V2")))

# Creating a graph 
g_edge_list <- graph_from_data_frame(edge_list, directed = FALSE)

# Adjacency matrix
g_adj_m <- as_adjacency_matrix(g_edge_list)

# Stochastic adjacency matrix
g_st_adj_m <- stochastic_matrix(g_edge_list, sparse = FALSE)

# Grph Laplacian 
g_laplacian <- laplacian_matrix(g_edge_list)

```

## 2. Given this [bipartite network](https://www.networkatlas.eu/exercises/5/2/data.txt), calculate the stochastic adjacency matrix of its projection. Project along the axis of size 248. (Note: don’t ignore the weights) 

### Python code from [book solutions](https://www.networkatlas.eu/exercise.htm)

```{python exercise 2, echo=FALSE}
# Load the data
G = nx.read_edgelist("../Data/ch5/bipartite_net_ch5.txt", data = [("weight", float),])

# Identify the bipartite sets
nodes = nx.algorithms.bipartite.basic.sets(G)

# Non-square adjacency matrix
adjmat = nx.algorithms.bipartite.matrix.biadjacency_matrix(G, nodes[0])

# Project along smaller axis
if adjmat.shape[0] == 248:
   adjmat_proj = adjmat.dot(adjmat.T)
else:
   adjmat_proj = adjmat.T.dot(adjmat)
   
# Make stocastic
adjmat_proj_stoc = adjmat_proj / adjmat_proj.sum(axis = 1)
print(adjmat_proj_stoc)

```

### R code

```{r excercise 2}
# Loading data 
edge_list <- data.frame(read.table("../Data/ch5/bipartite_net_ch5.txt", col.names = c("V1", "V2", "weight")))

# Creating a graph 
bib_g <- graph_from_data_frame(edge_list, directed = FALSE)

# Assigning corresponding type for all nodes 
for(i in 1:length(V(bib_g))) {
   if(grepl("a", V(bib_g)[i]$name)){
      V(bib_g)[i]$"type" <- "TRUE"
   }else{
      V(bib_g)[i]$"type" <- "FALSE"
   }
}

# Extracting the incidence matrix of the bip graph
adj_m_bip_g <- as_incidence_matrix(bib_g, attr = "weight", sparse = TRUE)

adjmat_proj <- t(adj_m_bip_g) %*% adj_m_bip_g

adjmat_proj_stoc = adjmat_proj / rowSums(adjmat_proj)

```

## 3. Calculate the eigenvalues and the right and left eigenvectors of the stochastic adjacency of this [bipartite network](https://www.networkatlas.eu/exercises/5/2/data.txt), using the same procedure applied in Exercise 5.2. Make sure to sort the eigenvalues in descending order (and sort the eigenvectors accordingly). Only take the real part of eigenvalues and eigenvectors, ignoring the imaginary part.

### Python code from [book solutions](https://www.networkatlas.eu/exercise.htm)

```{python exercise 3, echo=FALSE}
# Eigenvalues and Eigenvectors. Right eigenvectors first.
values, vectors_r = np.linalg.eig(adjmat_proj_stoc)

# We need to sort the eigenvalues and eigenvectors, since numpy returns them in random order
sorted_index = values.argsort()[::-1]
values = np.real(values[sorted_index])
vectors_r = np.real(vectors_r[:,sorted_index])
print(values)
print(vectors_r)

# Now left eigenvectors.
values, vectors_l = np.linalg.eig(adjmat_proj_stoc.T)
sorted_index = values.argsort()[::-1]
values = np.real(values[sorted_index])
vectors_l = np.real(vectors_l[:,sorted_index])
print(vectors_l)

```

### R code

```{r excercise 3}
# Extract eigen values and eigen vectors 

e_r <- eigen(adjmat_proj_stoc)

e_l <- eigen(t(adjmat_proj_stoc))

eigen_values <- e_r$values

eigen_vec_r <- e_r$vectors

eigen_vec_l <- e_l$vectors

```


