---
title: "Chapter6"
author: "SebastiánAyala"
date: "28/09/2021"
output: html_document
---

```{r libraries and python version, echo=FALSE, include=FALSE}
library(igraph)
library(ggplot2)
library(reticulate)
library(readr)
library(dplyr)
library(scales)
```

```{python packages, echo=FALSE}
# Import Python packages 
import networkx as nx
import matplotlib.pyplot as plt
import pandas as pd
import numpy as np 
from collections import Counter
from scipy.stats import linregress
import powerlaw as pl
from scipy.optimize import curve_fit
```
# Exercises chapter 6

## 4. Plot the degree distribution of this [network](https://www.networkatlas.eu/exercises/6/4/data.txt). Start from a plain degree distribution, then in log-log scale, finally plot the complement of the cumulative distribution.

### Python code from [book solutions](https://www.networkatlas.eu/exercise.htm)

```{python exercise 4, echo=FALSE}
# Load the data
G = nx.read_edgelist("../Data/ch6/edge_list_ch6.txt", delimiter = "\t")

# G.degrees return the degree dictionary, of which we only take the values and store
# them in a list. Counter will tell us how many elements in a list have a specific
# value (i.e. how many nodes have a specific degree). Then we convert it into a
# data frame for easier processing.
dd = Counter(dict(G.degree).values())
dd = pd.DataFrame(list(dd.items()), columns = ("k", "count")).sort_values(by = "k")

# We use the standard pandas plotting function to see the degree distribution and
# save it to file.
dd.plot(kind = "scatter", x = "k", y = "count", color = "#e41a1c")
plt.show()

# The standard pandas plotting function allows us to use a log-log scale via a
# parameter, then we again save it to file.
dd.plot(kind = "scatter", x = "k", y = "count", color = "#e41a1c", loglog = True)
plt.show()

# To make the CCDF we need to know how many nodes have degree equal to or higher
# than a specific value. So we sort the dataframe in descending degree order, so
# that the pandas cumsum function will calculate that for us. Then we normalize by
# the total degree sum, so that the count becomes a probability. We then sort in
# ascending degree value, to respect the convention.
ccdf = dd.sort_values(by = "k", ascending = False)
ccdf["cumsum"] = ccdf["count"].cumsum()
ccdf["ccdf"] = ccdf["cumsum"] / ccdf["count"].sum()
ccdf = ccdf[["k", "ccdf"]].sort_values(by = "k")

# Plot as usual and save it for later, since it's very pretty.
ccdf.plot(kind = "line", x = "k", y = "ccdf", color = "#e41a1c", loglog = True)
plt.show()

```

### R code

```{r excercise 4}
# Load edge list 
edge_list <- data.frame(read.table("../Data/ch6/edge_list_ch6.txt", col.names = c("V1", "V2")))

# Creating a graph 
g_edge_list <- graph_from_data_frame(edge_list, directed = FALSE)

deg_distrib <- function(network){
  deg_g <- as.numeric(igraph::degree(network))
  deg_g_freq <- as.data.frame(table(deg_g))
  deg_g_freq$deg_g <- as.numeric(as.character(deg_g_freq$deg_g))
  deg_g_freq
}

g_deg_distrib <- deg_distrib(g_edge_list)

colnames(g_deg_distrib) <- c("deg", "freq")

ggplot(g_deg_distrib, aes(x = deg, y = freq )) +
  theme_bw()+
  geom_point(size = 0.9)+
  xlab("k") +
  ylab("Count")

ggplot(g_deg_distrib, aes(x = deg, y = freq )) +
  theme_bw()+
  geom_point(size = 0.9)+
  xlab("k") +
  ylab("Count")+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))+
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))

# get the ecdf - Empirical Cumulative Distribution Function of freq
my_ecdf <- ecdf(g_deg_distrib$deg)

# now put the ecdf and its complementary in a data.frame
g_deg_distrib_ccdf <- data.frame(x = sort(g_deg_distrib$deg),
                    y = 1-my_ecdf(sort(g_deg_distrib$deg) ))

ggplot(g_deg_distrib_ccdf, aes(x, y)) +
  theme_bw()+
  geom_line()+
  xlab("x") +
  ylab("p(k>=x)")+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))+
  scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))

```

## 5. Estimate the power law exponent of the CCDF degree distribution from Exercise 6.4. First by a linear regression on the log-log plane, then by using the powerlaw package. Do they agree? Is this a shifted power law? If so, what's k min? (Hint: powerlaw can calculate this for you)

### Python code 

```{python exercise 5, echo=FALSE}
# We take the logarithm in base 10 of both degree and CCDF. Then we simply do a linear regression. The slope is
# the exponent. The intercept needs to be the power of 10, to undo the logarithm operation. Look at that
# r-squared!
logcdf = np.log10(ccdf[["k", "ccdf"]])
slope, log10intercept, r_value, p_value, std_err = linregress(logcdf["k"], logcdf["ccdf"])
print("CCDF Fit: %1.4f x ^ %1.4f (R2 = %1.4f, p = %1.4f)" % (10 ** log10intercept, slope, r_value ** 2, p_value))

# With the powerlaw package, fitting the CCDf is simple. It will store results in the .power_law property. To
# get the actual k_min, we need to find the degree value corresponding to the probability in .power_law.xmin:
# pandas makes it easy. This is definitely a shifted power law. (Kappa contains the intercept information)
results = pl.Fit(ccdf["ccdf"])
k_min = ccdf[ccdf["ccdf"] == results.power_law.xmin]["k"]
print("Powerlaw CCDF Fit: %1.4f x ^ -%1.4f (k_min = %d)" % (10 ** results.power_law.Kappa, results.power_law.alpha, k_min))

# Let's plot the best fit.
ccdf["fit"] = (10 ** results.power_law.Kappa) * (ccdf["k"] ** -results.power_law.alpha)
ax = plt.gca()
ccdf.plot(kind = "line", x = "k", y = "ccdf", color = "#e41a1c", loglog = True, ax = ax)
ccdf.plot(kind = "line", x = "k", y = "fit", color = "#377eb8", loglog = True, ax = ax)
plt.show()

```

### R code

```{r excercise 5}
logccdf <- as.data.frame(log10(g_deg_distrib_ccdf[1:474,]))

deg_distrib_lm <- lm(x ~ y, data = logccdf)

summary(deg_distrib_lm)

```

## 6. Find a way to fit the truncated power law of this [network](https://www.networkatlas.eu/exercises/6/6/data.net). Hint: use the scipy.optimize.curve_fit to fit an arbitrary function and use the functional form I provide in Chapter 6 of the book

### Python code 

```{python exercise 6, echo=FALSE}

# Load the data and redo the CCDF.
G = nx.read_pajek("../Data/ch6/pajek_net_ch6.txt")
dd = Counter(dict(G.degree).values())
dd = pd.DataFrame(list(dd.items()), columns = ("k", "count")).sort_values(by = "k")
ccdf = dd.sort_values(by = "k", ascending = False)
ccdf["cumsum"] = ccdf["count"].cumsum()
ccdf["ccdf"] = ccdf["cumsum"] / ccdf["count"].sum()
ccdf = ccdf[["k", "ccdf"]].sort_values(by = "k")

# Let's define a custom function which is a power law with its exponential truncation. We
# also define its logarithm, because we fit it to the log of the CCDF in curve_fit. This
# is done because we want to minimize the relative error, not the absolute error (since
# the tail of the distribution is very important, but it contributes very little to the
# absolute error). Then we plot.
def f(x, a, l):
   return (x ** a) * np.exp(-l * x) 

def log_f(x, a, l):
   return np.log10(f(x, a, l))

popt, pcov = curve_fit(log_f, ccdf["k"], np.log10(ccdf["ccdf"]), p0 = (1, 1))
ccdf["fit"] = ccdf.apply(lambda x: f(x["k"], popt[0], popt[1]), axis = 1)

ax = plt.gca()
ccdf.plot(kind = "line", x = "k", y = "ccdf", color = "#e41a1c", loglog = True, ax = ax)
ccdf.plot(kind = "line", x = "k", y = "fit", color = "#377eb8", loglog = True, ax = ax)

plt.show()
```



