---
title: "Chapter4"
author: "SebastiánAyala"
date: "8/24/2021"
output: html_document
---

```{r libraries and python version, echo=FALSE, include=FALSE}
library(igraph)
library(HyperG)
library(ggplot2)
library(reticulate)

# Specifying which version of python to use.
use_python("/home/asayalar/anaconda3/envs/networks/bin/python3.8", 
           required = T)

```

```{python packages, echo=FALSE}
# Import Python packages 
import networkx as nx
import matplotlib.pyplot as plt
import pandas as pd
from collections import defaultdict
from itertools import combinations
```


# Exercises chapter 4

## 1. The network in http://www.networkatlas.eu/exercises/4/1/
data.txt is bipartite. Identify the nodes in either type and find the
nodes, in either type, with the most neighbors

### Python code 

```{python exercise 1, echo=FALSE}

# Load the data
G = nx.read_edgelist("Data/bipartite_net_ch4.txt")

# Divide the nodes in the two bipartite sets
nodes = nx.algorithms.bipartite.basic.sets(G)

# Find the node in type 0 with the most neighbors
node_neighbors = {n: len(set(G.neighbors(n))) for n in nodes[0]}
maxnode = max(node_neighbors, key = node_neighbors.get)
print(maxnode, node_neighbors[maxnode])

# Find the node in type 1 with the most neighbors
node_neighbors = {n: len(set(G.neighbors(n))) for n in nodes[1]}
maxnode = max(node_neighbors, key = node_neighbors.get)
print(maxnode, node_neighbors[maxnode])
```

### R code

```{r excercise 1}
# Load edge list 
edge_list <- data.frame(read.table("../Data/bipartite_net_ch4.txt"))

# Create the bipartite graph 
bip_g <- graph.empty(directed = F)
node.out <- unique(edge_list$V1) #stringsAsFactor = F in data frame
node.in <- unique(edge_list$V2) #stringsAsFactor = F in data frame
bip_g <- add.vertices(bip_g,nv=length(node.out),attr=list(name=node.out),type=rep(FALSE,length(node.out)))
bip_g <- add.vertices(bip_g,nv=length(node.in),attr=list(name=node.in),type=rep(TRUE,length(node.in)))
edge.list.vec <- as.vector(t(as.matrix(data.frame(edge_list))))
bip_g <- add.edges(bip_g,edge.list.vec)

# Count the number of neighbors 
set1 <- data.frame(node = node.in, n_neighbors = ego_size(bip_g, order = 1, node.in))

max_set1 <- set1[which.max(set1$n_neighbors),]

set2 <- data.frame(node = node.out, n_neighbors = ego_size(bip_g, order = 1, node.out))

max_set2 <- set2[which.max(set2$n_neighbors),]

```


## 2. This [network](https://www.networkatlas.eu/exercises/4/2/data.txt) is multilayer. The data has three columns: source and target node, and edge type. The edge type is either the numerical id of the layer, or "C" for an inter-layer coupling. Given that this is a one-to-one multilayer network, determine whether this network has a star, clique or chain coupling.

### Python code 

```{python exercise 2, echo=FALSE}

# Load the data. We need to import as multigraph, or networkx will collapse the edges.
# We also need to make sure to import the edge type information.
G = nx.read_edgelist("../Data/multilayer_net_ch4.txt", create_using = nx.MultiGraph(), data = [("edgetype", str),])

# Determine how many layers we have
layers = len(set(e[2]["edgetype"] for e in G.edges(data = True))) - 1

# If this is a chain coupling, all nodes can have at most two coupling edges
node_cedges = defaultdict(int)
for e in G.edges(data = True):
   if e[2]["edgetype"] == "C":
      node_cedges[e[0]] += 1
      node_cedges[e[1]] += 1

if max(node_cedges.values()) == 2:
   print("Chain coupling")
   
# If this is a clique coupling, all nodes must have more than one coupling edge
if min(node_cedges.values()) > 1:
   print("Clique coupling")

# If this is a star coupling, all nodes must have either one edge, or |L| - 1 coupling edges
if set([1, layers - 1]) == set(node_cedges.values()):
   print("Star coupling")

```

### R code

```{r excercise 2}
# Load dataframe 
multil_net_df <- data.frame(read.table("../Data/multilayer_net_ch4.txt", col.names = c("from", "to", "layer")))

# Create the multilayer graph 
multig <- graph_from_data_frame(multil_net_df, directed=FALSE)
multig

# Number of edges per layer
layers <- length(table(E(multig)$layer)) - 1

# df of nodes with coupling edges

df_ce_to <- data.frame(table(multil_net_df[multil_net_df$layer== "C", ]$to))

colnames(df_ce_to) <- c('node', 'n_coupling_edges')

df_ce_from <- data.frame(table(multil_net_df[multil_net_df$layer== "C", ]$from))

colnames(df_ce_from) <- c('node', 'n_coupling_edges')

df_ce = merge(x=df_ce_to,y=df_ce_from,by=c("node", 'n_coupling_edges'), all=TRUE)

# Check if the multilayer net is chain coupling, where all nodes can have at most two coupling edges

if(max(df_ce$n_coupling_edges) == 2 ){
    print("Chain coupling")
}

# Check if the multilayer net is clique coupling, where all nodes must have more than one coupling edge

if (min(df_ce$n_coupling_edges) > 1){
     print("Clique coupling")
}

# Check if the multilayer net is star coupling, where all nodes must have either one edge, or |L| - 1 coupling edges

if (df_ce$n_coupling_edges == 1 | df_ce$n_coupling_edges == layers-1){
    print("Star coupling")
}

```

## 3. This [network](https://www.networkatlas.eu/exercises/4/3/data.txt) is a hypergraph, with a hyperedge per line. Transform it in a unipartite network in which each hyperedge is split in edges connecting all nodes in the hyperedge. Then transform it into a bipartite network in which each hyperedge is a node of one type and its nodes connect to it.

### Python code 

```{python exercise 3, echo=FALSE}

# Load the data.
hyperedges = []
with open("../Data/hypergraph_net_ch4.txt", 'r') as f:
   for line in f:
      hyperedges.append(line.strip().split())

# Make it into a unipartite graph
G_uni = nx.Graph()
for hyperedge in hyperedges:
   for combination in combinations(hyperedge, 2):
      G_uni.add_edge(*combination)

# Make it into a bipartite graph
G_bip = nx.Graph()
hyperid = 1
for hyperedge in hyperedges:
   for node in hyperedge:
      G_bip.add_edge(node, "h%s" % hyperid)
   hyperid += 1
```

### R code

```{r excercise 3}

# Load hypergraph with HyperG package 

hyperg <- hypergraph_from_edgelist(list(c(1, 2, 3, 4, 5, 6), c(2,4,7,8), 
                                   c(4,5,6,7,10), c(6,9,10,11,12), c(11,13,14,15), 
                                   c(11,12,13), c(14,15,16,17,18), c(15,17,19,20), 
                                   c(21,22,23), c(23,24), c(24,25,26), 
                                   c(25,27,28,29)))

# Plot the hypergraph
#plot(hyperg)

# Convert hypergraph to unipartite graph 
g <- as.graph(hyperg)

# Convert hypergraph to bipartite graph

bipart_g <- as.bipartite(hyperg)
V(bipart_g)[type == TRUE]$color <- "orange"
V(bipart_g)[type == FALSE]$color <- "green"

# Plot the three graphs 
par(mfrow=c(1, 3))
plot(hyperg)
title("Hypergraph")
plot(bipart_g)
title("Bipartite graph")
plot(g)
title("Unipartite graph")
```


## 4. This [network](https://www.networkatlas.eu/exercises/4/4/data.txt) is dynamic, the third and fourth columns of the edge list tell you the first and last snapshot in which the edge was continuously present. An edge can reappear if the edge was present in two discontinuous time periods. Aggregate it using a disjoint window of size 3.

### Python code 

```{python exercise 4, echo=FALSE}

# Load the data. We need to import as multigraph, or networkx will collapse the edges.
# We also need to make sure to import the edge type information.
G = nx.read_edgelist("../Data/dynamic_net_ch4.txt", create_using = nx.MultiGraph(), data = [("start", int), ("end", int)])

# Since the window is disjoint and of size three, we need to group the 1-3, 4-6, and 7-9 snapshots.
first_window = nx.Graph()
second_window = nx.Graph()
third_window = nx.Graph()
for e in G.edges(data = True):
   if e[2]["start"] <= 3:
      first_window.add_edge(e[0], e[1])
   if e[2]["start"] <= 6 and e[2]["end"] > 3:
      second_window.add_edge(e[0], e[1])
   if e[2]["end"] > 6:
      third_window.add_edge(e[0], e[1])

```

### R code

```{r excercise 4}
# Load dataframe 
multil_net_df2 <- data.frame(read.table("../Data/dynamic_net_ch4.txt", col.names = c("from", "to", "start_snap", "end_snap")))

# Create the multilayer graph 
multig2 <- graph_from_data_frame(multil_net_df2, directed=FALSE)
multig2

# Create networks with disjoint windows of size 3
first_window <- make_empty_graph(n = 12, directed = FALSE)
second_window <- make_empty_graph(n = 12, directed = FALSE)
third_window <- make_empty_graph(n = 12, directed = FALSE)

first_window_elist <- c()
i <-1
second_window_elist <- c()
j <- 1
third_window_elist <- c()
z <- 1

for(e in length(E(multig2))){
    if(E(multig2)$start_snap[e] <= 3){
        first_window_elist[i:i+1] <- c(E(multig2)$start_snap[e], 
                                  E(multig2)$end_snap[e])
        i <- i + 2
    }
    if(E(multig2)$start_snap[e] <= 6 & E(multig2)$end_snap[e] > 3){
        second_window_elist[j:j+1] <- c(E(multig2)$start_snap[e], 
                                  E(multig2)$end_snap[e])
        j <- j + 2
    }
    
    if(E(multig2)$end_snap[e] > 6){
        third_window_elist[z:z+1] <- c(E(multig2)$start_snap[e], 
                                  E(multig2)$end_snap[e])
        z <- z + 2
    }
}
   
# Plot the three graphs 
# par(mfrow=c(1, 3))
# plot(first_window)
# title("First window")
# plot(second_window)
# title("Second window")
# plot(third_window)
# title("Third window")

first_window_elist
second_window_elist
third_window_elist

```